---
title: "STAA-566: Data Vis"
subtitle: "Assignment 3: Maps"
author: "Penny Grant"
date: "`r Sys.Date()`"
header-includes:
  - \setmainfont{Noto Sans}
  - \setmonofont{Noto Sans Mono}
  - \usepackage{physics}
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load packages

```{r, message = FALSE}
require(tidyverse)
require(ggplot2)
require(sf)
require(rworldmap)
require(ggthemes)
require(viridis)
require(plotly)
```

# Import Dataset

This data set was published by the World Happiness Report and downloaded from [kaggle](https://www.kaggle.com/datasets/mathurinache/world-happiness-report). It uses information from the Gallup World Survey and bases scores off several different factors and the degree to which they contribute to the country's happiness.

```{r,warning=FALSE, message = FALSE}
happiness <- read_csv("2022happinessReport.csv")
```

Now I convert my data to include Country codes.

```{r}
world_sf <- st_as_sf(getMap(resolution = "low")) %>% 
  left_join(happiness, by = c("GEOUNIT"="Country"))
```

# Plot Data

First I draw the world map on which to include my data. Then I add coloring by the happiness score. I also created a text field to be able to have a tooltip with the happiness score and country's ranking.

```{r, warning=FALSE}
crs_robin <- "+proj=robin +lat_0=0 +lon_0=0 +x0=0 +y0=0"

h_wm <- ggplot()+
  geom_sf(data=world_sf,
          mapping = aes(fill=`Happiness score`,
                        text=paste("Happiness Ranking:",RANK,"<br>Country:",
                                  GEOUNIT,"<br>Happiness Score:",`Happiness score`)),
          col="white", size =.1)
```

Then I adjust the themes and color schemes

```{r}
h_wm <- h_wm +
  theme_classic(base_size = 12,base_family = "sans")+
  coord_sf(crs=crs_robin) +
  scale_fill_viridis() +
  theme(plot.title = element_text(hjust=.5),legend.position = "none")
```

Finally, I add a title and an interactive element
```{r}
h_wm <- h_wm + ggtitle("2022 Happiness Score")


# add tooltips
h_wm <- ggplotly(h_wm,tooltip= "text")
```

The tooltips show where the country ranks as far as it's happiness score, its name, and the happiness score itself.

`r h_wm`


